---
title: "R Notebook"
output: html_notebook
---

# load library
```{r}
options (scipen = 999)
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(vip)
library(parallel)
library(doParallel)
library(embed)
library(rpart.plot) 
library(rpart) 
library(kableExtra)
library(GGally)
library(fastshap)  
library(MASS)
library(caret)
library(glmnet)
library(reshape2)
library(textrecipes)
library(DALEX)    # new
library(DALEXtra) # new
library(solitude)
```

# import data
```{r}
loan <- read_csv("loan_train.csv") %>%
  clean_names() 
head(loan)
```
# skim 
```{r}
loan %>%
  skim() ->skim
```
# Data Munging
```{r}
loan %>%
  dplyr::select(-mths_since_last_record,-mths_since_last_delinq,-next_pymnt_d,-url,-emp_title)%>%
  mutate(term = as.numeric(chartr('months','      ',term))) %>%
  mutate(int_rate = as.numeric(chartr('%',' ',int_rate))) %>%
  mutate(sub_grade = as.numeric(substr(sub_grade,2,3))) %>%
  mutate(zip_code = as.numeric(chartr('xx','  ',zip_code))) %>%
  mutate(revol_util = as.numeric(chartr('%',' ',revol_util))) %>%
  mutate(earliest_cr_line_mth = match(tolower(substr(earliest_cr_line,1,3)), tolower(month.abb))) %>%
  mutate(earliest_cr_line_yr = as.numeric(substr(earliest_cr_line,5,9))) %>%
  mutate(issue_d_mth = match(tolower(substr(issue_d,1,3)), tolower(month.abb))) %>%
  mutate(issue_d_yr = as.numeric(substr(issue_d,5,9))) %>%
  mutate(last_pymnt_d_mth = match(tolower(substr(last_pymnt_d,1,3)), tolower(month.abb))) %>%
  mutate(last_pymnt_d_yr = as.numeric(substr(last_pymnt_d,5,9))) %>%
  mutate(last_credit_pull_d_mth = match(tolower(substr(last_credit_pull_d,1,3)), tolower(month.abb))) %>%
  mutate(last_credit_pull_d_yr = as.numeric(substr(last_credit_pull_d,5,9))) %>%
  dplyr::select(-issue_d,-earliest_cr_line,-last_pymnt_d,-last_credit_pull_d,
                -collections_12_mths_ex_med,-policy_code,-chargeoff_within_12_mths)->loan

addr_state_freq <- loan %>%
  group_by(addr_state) %>%
  summarise(addr_state_freq = n()) 
loan <- loan %>%
  left_join(addr_state_freq)
loan %>%
  dplyr::select(-addr_state) ->loan


title_freq <- loan %>%
  group_by(title) %>%
  summarise(title_freq = n()) 
loan <- loan %>%
  left_join(title_freq)
loan %>%
  dplyr::select(-title) ->loan

loan%>%
  mutate(issue_d=issue_d_yr+issue_d_mth/12,
         earliest_cr_line=earliest_cr_line_yr+earliest_cr_line_mth/12,
         last_pymnt_d=last_pymnt_d_yr+last_pymnt_d_mth/12,
         last_credit_pull_d=last_credit_pull_d_yr+last_credit_pull_d_mth/12)->loan


        
```
# Dealing with Text using Sentiment
```{r}
library(tidytext)
afinn <- get_sentiments("afinn")

sent_desc <- loan %>% 
  unnest_tokens(word, desc) %>% 
  filter(!word %in% stop_words) %>% 
  inner_join(afinn) %>% 
  group_by(id) %>% 
  summarise(sentiment_desc = sum(value))

loan <- loan %>% 
  left_join(sent_desc)

```

# Target Encoding & As Factors
```{r}
loan %>%
  mutate(loan_status = as.factor(if_else(loan_status == "default",1,0)))%>%
  mutate_if(is.character,as_factor) %>%
  mutate(desc=as.character(desc))->loan
```

# Data Exploration
## Explore target
```{r}
loan %>%
  count(loan_status) %>%
  mutate(pct = n/sum(n)) -> default

default %>%
  ggplot(aes(x=loan_status, y=pct)) +
  geom_col() +
  geom_text(aes(label=pct) ,color="red") + 
  labs(title="Default Rate")
```
## Explore numerics
```{r}
boxplot<- function(a,b){
  ggplot(loan, aes(x={{a}},y=loan_status,group=loan_status)) + geom_boxplot() +labs(title = b)
}

boxplot(member_id,"member_id")
boxplot(loan_amnt,"loan_amnt")
boxplot(funded_amnt,"funded_amnt")
boxplot(funded_amnt_inv,"funded_amnt_inv")
boxplot(term,"term")
boxplot(int_rate,"int_rate")
boxplot(installment,"installment")
boxplot(sub_grade,"sub_grade")
boxplot(annual_inc,"annual_inc")
boxplot(zip_code,"zip_code")
boxplot(dti,"dti")
boxplot(delinq_2yrs,"delinq_2yrs")
boxplot(fico_range_low,"fico_range_low")
boxplot(fico_range_high,"fico_range_high")
boxplot(inq_last_6mths,"inq_last_6mths")
boxplot(open_acc,"open_acc")
boxplot(pub_rec,"pub_rec")
boxplot(revol_bal,"revol_bal")
boxplot(revol_util,"revol_util")
boxplot(total_acc,"total_acc")
boxplot(out_prncp,"out_prncp")
boxplot(out_prncp_inv,"out_prncp_inv")
boxplot(total_rec_late_fee,"total_rec_late_fee")
boxplot(last_pymnt_amnt,"last_pymnt_amnt")
boxplot(acc_now_delinq,"acc_now_delinq")
boxplot(delinq_amnt,"delinq_amnt")
boxplot(pub_rec_bankruptcies,"pub_rec_bankruptcies")
boxplot(tax_liens,"tax_liens")
boxplot(earliest_cr_line_mth,"earliest_cr_line_mth")
boxplot(earliest_cr_line_yr,"earliest_cr_line_yr")
boxplot(issue_d_mth,"issue_d_mth")
boxplot(issue_d_yr,"issue_d_yr")
boxplot(last_pymnt_d_mth,"last_pymnt_d_mth")
boxplot(last_pymnt_d_yr,"last_pymnt_d_yr")
boxplot(last_credit_pull_d_mth,"last_credit_pull_d_mth")
boxplot(last_credit_pull_d_yr,"last_credit_pull_d_yr")
boxplot(addr_state_freq,"addr_state_freq")
```

## Explore characters
### barplot
```{r}
bar_plot <- function(a,colname){
    loan %>%
    ggplot(., aes(a)) + 
    geom_bar(aes(fill = loan_status), position = "fill") +
    labs(title = paste("Barplot for",colname), x = colname, y="percent")+
    geom_hline(yintercept = 0.15035)+
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5))
}

bar_plot(loan$grade,"grade")
bar_plot(loan$emp_length,"emp_length")
bar_plot(loan$home_ownership,"home_ownership")
bar_plot(loan$verification_status,"verification_status")
bar_plot(loan$pymnt_plan,"pymnt_plan")
bar_plot(loan$purpose,"purpose")
bar_plot(loan$application_type,"application_type")
```

## Explore correlation
```{r}
library(corrplot)
cormat <- cor(select_if(loan%>%na.omit(),is.numeric))
round(cormat, 2)
corrplot(cormat,tl.cex  = 0.5)
```


# define train/test split 
```{r}
set.seed(42)
train_test_spit<- initial_split(loan, prop = 0.7)

train <- training(train_test_spit)
test  <- testing(train_test_spit)

train_cv_folds <- vfold_cv(train, v=5)
```

# Define Recipe
```{r}
loan_recipe <- recipe(loan_status ~ . , data = train) %>%
  step_novel(all_nominal(), -all_outcomes())%>%
  step_unknown(all_nominal(), -all_outcomes()) %>%
  step_meanimpute(all_numeric())%>%
  
  step_impute_mode(all_nominal_predictors()) %>%
  step_tokenize(desc) %>%
  step_tokenfilter(desc, max_tokens = 100) %>%
  step_stopwords(desc) %>%
  step_tfidf(desc) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_rm(id)
  
loan_recipe
```
# logistic
```{r}
recipe <- recipe(loan_status ~ . , data = train) %>%
  step_novel(all_nominal(), -all_outcomes())%>%
  step_unknown(all_nominal(), -all_outcomes()) %>%
  step_meanimpute(all_numeric())%>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_tokenize(desc) %>%
  step_tokenfilter(desc, max_tokens = 100) %>%
  step_stopwords(desc) %>%
  step_tfidf(desc) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_rm(id,funded_amnt,int_rate,dti,addr_state_freq,delinq_2yrs,out_prncp,acc_now_delinq,delinq_amnt,pub_rec_bankruptcies,tfidf_desc_08,tfidf_desc_09,tfidf_desc_10)
  

logistic_spec <- logistic_reg() %>%
  set_mode("classification") %>%
  set_engine("glm")

logistic_full_wf1 <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(logistic_spec) %>%
  fit(train)

logistic_full_wf1 %>%
  pull_workflow_fit() %>%
  tidy() %>%
  mutate(across(is.numeric,round,3))

logistic_full_wf1 %>%
  pull_workflow_fit() %>%
  vip(10)

options(yardstick.event_first = FALSE)
predict(logistic_full_wf1, train, type="prob") %>%
  bind_cols(predict(logistic_full_wf1, train, type="class")) %>%
  bind_cols(train) ->logistic_train1

predict(logistic_full_wf1, test, type="prob") %>%
  bind_cols(predict(logistic_full_wf1, test, type="class")) %>%
  bind_cols(test) -> logistic_test1


logistic_train1 %>% 
    metrics(loan_status, .pred_1, estimate = .pred_class) %>%
    mutate(part="training") %>%
    bind_rows( logistic_test1 %>% 
                 metrics(loan_status, .pred_1, estimate = .pred_class) %>%
                 mutate(part="testing") ) %>%
    filter(.metric %in% c('accuracy','roc_auc','mn_log_loss')) %>%
    pivot_wider(names_from = .metric, values_from=.estimate)
  
logistic_train1 %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
logistic_test1 %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
  # -- ROC Charts 
  logistic_train1 %>%
  mutate(model = "train") %>%
  bind_rows(logistic_test1 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(loan_status, .pred_1) %>%
  autoplot() 

```
## KS and ROC in operating range
```{r}
# -- calculate KS  -- 
logistic_test1 %>%
 roc_curve(loan_status, .pred_1) %>%
  mutate(fpr = round((1 - specificity),2),
         tpr = round(sensitivity,3),
         score_threshold = round(.threshold,3)) %>%
  mutate(diff_tprfpr = tpr - fpr) %>%
  slice_max(diff_tprfpr,n=1, with_ties = FALSE) %>%
  dplyr::select(fpr,tpr,score_threshold,ks = diff_tprfpr)

logistic_test1 %>%
 roc_curve(loan_status, .pred_1) %>%
  autoplot() +
  geom_vline(aes(xintercept=0.06, color="red")) +
  labs(title="ROC operating at 6% FPR")

logistic_test1 %>%
  pr_curve(loan_status, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3),
    f1 = round(2*precision*recall/(precision+recall),3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold),
            f1 = max(f1))
logistic_test1 %>%
  ggplot(aes(.pred_1, fill=loan_status)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.11)) +
  labs(title="logistic score distribution")

```

# Random Forest

## tune grid
```{r}
rf_tune <- rand_forest(
    trees  = tune(),
    min_n = tune()) %>% 
    set_engine("ranger", importance = "impurity") %>% 
    set_mode("classification")

# -- setup workflow 
rf_workflow <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_tune) 

# -- setup your tuning grid -- random force 
tune_grid1 <- grid_random(trees(),
                         min_n(),
                        size = 5)
print(tune_grid1)

# -- setup parallel process 
all_cores <- detectCores(logical = TRUE)
sprintf("# of Logical Cores: %d", all_cores)
cl <- makeCluster(all_cores)
registerDoParallel(cl)

# -- train!! K times for each parameter -- 
rf_tuning_results <- rf_workflow %>% 
  tune_grid(
    resamples = train_cv_folds,
    grid = tune_grid1,
    control = control_resamples(save_pred = TRUE)
    )

rf_tuning_results
```
## refit
```{r}
rf_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err))

rf_best <- rf_tuning_results %>%
  select_best("roc_auc") 
print(rf_best)

rf_final_wf <- rf_workflow %>% 
  finalize_workflow(rf_best)
print(rf_final_wf)

rf_final_fit  <- rf_final_wf %>%
  fit(data = train) 
```
## evaluation
```{r}
options(yardstick.event_first = FALSE)
predict(rf_final_fit, train, type="prob") %>%
  bind_cols(predict(rf_final_fit, train, type="class")) %>%
  bind_cols(train) ->rf_train2

predict(rf_final_fit, test, type="prob") %>%
  bind_cols(predict(rf_final_fit, test, type="class")) %>%
  bind_cols(test) -> rf_test2


rf_train2 %>% 
    metrics(loan_status, .pred_1, estimate = .pred_class) %>%
    mutate(part="training") %>%
    bind_rows( rf_test2 %>% 
                 metrics(loan_status, .pred_1, estimate = .pred_class) %>%
                 mutate(part="testing") ) %>%
    filter(.metric %in% c('accuracy','roc_auc','mn_log_loss')) %>%
    pivot_wider(names_from = .metric, values_from=.estimate)
  
  rf_train2 %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
   rf_test2 %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
  # -- ROC Charts 
  rf_train2 %>%
  mutate(model = "train") %>%
  bind_rows(rf_test2 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(loan_status, .pred_1) %>%
  autoplot() 

    # -- variable importance: top 10
  rf_final_fit %>%
    pull_workflow_fit() %>%
  vip(num_features = 10)
```

## handy
```{r}
rf_model <- rand_forest(
    trees  = 199,min_n=18) %>% 
    set_engine("ranger", importance = "impurity") %>% 
    set_mode("classification")
rf_workflow <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_model) %>%
  fit(data=train)

options(yardstick.event_first = FALSE)
predict(rf_workflow, train, type="prob") %>%
  bind_cols(predict(rf_workflow, train, type="class")) %>%
  bind_cols(train) ->rf_train1

rf_train1 %>%
  metrics(loan_status, estimate = .pred_class, .pred_1)

predict(rf_workflow, test, type="prob") %>%
  bind_cols(predict(rf_workflow, test, type="class")) %>%
  bind_cols(test) -> rf_test1

rf_test1 %>%
  metrics(loan_status, estimate = .pred_class, .pred_1)

rf_train1 %>% 
    metrics(loan_status, .pred_1, estimate = .pred_class) %>%
    mutate(part="training") %>%
    bind_rows( rf_test1 %>% 
                 metrics(loan_status, .pred_1, estimate = .pred_class) %>%
                 mutate(part="testing") ) %>%
    filter(.metric %in% c('accuracy','roc_auc','mn_log_loss')) %>%
    pivot_wider(names_from = .metric, values_from=.estimate)
```
## KS and ROC in operating range
```{r}
# -- calculate KS  -- 
rf_test1 %>%
 roc_curve(loan_status, .pred_1) %>%
  mutate(fpr = round((1 - specificity),2),
         tpr = round(sensitivity,3),
         score_threshold = round(.threshold,3)) %>%
  mutate(diff_tprfpr = tpr - fpr) %>%
  slice_max(diff_tprfpr,n=1, with_ties = FALSE) %>%
  dplyr::select(fpr,tpr,score_threshold,ks = diff_tprfpr)

rf_test1 %>%
 roc_curve(loan_status, .pred_1) %>%
  autoplot() +
  geom_vline(aes(xintercept=0.05, color="red")) +
  labs(title="ROC operating at 5% FPR")

rf_test1 %>%
  pr_curve(loan_status, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3),
    f1 = round(2*precision*recall/(precision+recall),3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold),
            f1 = max(f1))

rf_test1 %>%
  ggplot(aes(.pred_1, fill=loan_status)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.255)) +
  labs(title="rf score distribution")
```

# XGB
## tune grid
```{r}
xgb_model <- boost_tree(
  trees = 200, 
  tree_depth = 10,       ## how deep of a tree, model complexity
  min_n = tune(),            ## minimum number of observations 
  #learn_rate = tune()        
) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

xgb_model

# -- setup workflow 
xgb_workflow <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(xgb_model) 
# -- setup your tuning grid -- random force 
tune_grid <- grid_random(#trees(),
                         min_n(),
                         #tree_depth(),
                        #learn_rate(),
                        
                          size = 5)
print(tune_grid)
```
## Review Tuning Results 
```{r}
 xgb_tuning_results <- xgb_workflow %>% 
   tune_grid(
     resamples = train_cv_folds,
     grid = tune_grid,
     control = control_resamples(save_pred = TRUE)
     )
 
 xgb_tuning_results

## -- results of tuning -- 
 xgb_tuning_results %>% 
   collect_metrics() %>%
   mutate_if(is.numeric, round,3) %>% 
   pivot_wider(names_from = .metric, values_from=c(mean, std_err))
```
## selecting "best" parameters 
```{r}
xgb_tuning_results %>%
  show_best("roc_auc") %>%
  print()

xgb_best <- xgb_tuning_results %>%
  select_best("roc_auc") 

print(xgb_best)
```
## refitting workflow with "best" parameter
```{r}
library(glmnet)
library(workflows)
xgb_final_wf <- xgb_workflow %>% 
  finalize_workflow(xgb_best)

print(xgb_final_wf)

xgb_final_fit  <- xgb_final_wf %>%
  fit(data = train) 
```

## score xgb model
```{r}
predict(xgb_final_fit , train, type="prob") %>%
    bind_cols(predict(xgb_final_fit, train, type="class")) %>%
    bind_cols(.,train)-> scored_xgb_train 

predict(xgb_final_fit , test, type="prob") %>%
    bind_cols(predict(xgb_final_fit, test, type="class")) %>%
    bind_cols(.,test) -> scored_xgb_test   

scored_xgb_train %>% 
    metrics(loan_status, .pred_1, estimate = .pred_class) %>%
    mutate(part="training") %>%
    bind_rows( scored_xgb_test %>% 
                 metrics(loan_status, .pred_1, estimate = .pred_class) %>%
                 mutate(part="testing") ) %>%
    filter(.metric %in% c('accuracy','roc_auc','mn_log_loss')) %>%
    pivot_wider(names_from = .metric, values_from=.estimate)
  
scored_xgb_train %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
scored_xgb_test %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
  # -- ROC Charts 
scored_xgb_train %>%
  mutate(model = "train") %>%
  bind_rows(scored_xgb_test %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(loan_status, .pred_1) %>%
  autoplot() 

    # -- variable importance: top 10
  xgb_final_fit %>%
    pull_workflow_fit() %>%
  vip(num_features = 10)
```

## handy model
```{r}
xgb_model <- boost_tree(
    trees = 200,  
    min_n=14,
    tree_depth=10,
    #learn_rate= 0.012312104
    ) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

# -- define workflows
xgb_model_wf <- workflow() %>%
  add_recipe(loan_recipe) %>%
  add_model(xgb_model) 

xgb_model_fit <- xgb_model_wf %>%
            fit(data=train)
```
## score
```{r}
options(yardstick.event_first = FALSE)
scored_train <-predict(xgb_model_fit  , train, type="prob") %>%
bind_cols(predict(xgb_model_fit  , train, type="class")) %>%
bind_cols(train)

 scored_test <-predict(xgb_model_fit  , test, type="prob") %>%
bind_cols(predict(xgb_model_fit , test, type="class")) %>%
bind_cols(test)

scored_train %>% 
    metrics(loan_status, .pred_1, estimate = .pred_class) %>%
    mutate(part="training") %>%
    bind_rows( scored_test %>% 
                 metrics(loan_status, .pred_1, estimate = .pred_class) %>%
                 mutate(part="testing") ) %>%
    filter(.metric %in% c('accuracy','roc_auc')) %>%
    pivot_wider(names_from = .metric, values_from=.estimate)

options(yardstick.event_first = FALSE)
scored_train %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
scored_test %>%
    conf_mat(loan_status, .pred_class) %>%
    autoplot(type = "heatmap")
  
  # -- ROC Charts 
scored_train %>%
  mutate(model = "train") %>%
  bind_rows(scored_test %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(loan_status, .pred_1) %>%
  autoplot() 

xgb_model_fit %>%
    pull_workflow_fit() %>%
  vip(num_features = 40)

recipe <- recipe(loan_status ~ . , data = train) %>%
  step_novel(all_nominal(), -all_outcomes())%>%
  step_unknown(all_nominal(), -all_outcomes()) %>%
  step_meanimpute(all_numeric())%>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_tokenize(desc) %>%
  step_tokenfilter(desc, max_tokens = 100) %>%
  step_stopwords(desc) %>%
  step_tfidf(desc) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_rm(id,funded_amnt,int_rate,delinq_2yrs,out_prncp,acc_now_delinq,delinq_amnt,pub_rec_bankruptcies,tfidf_desc_08,tfidf_desc_09,tfidf_desc_10)
```
## KS and ROC in operating range
```{r}
# -- calculate KS  -- 
scored_test %>%
 roc_curve(loan_status, .pred_1) %>%
  mutate(fpr = round((1 - specificity),2),
         tpr = round(sensitivity,3),
         score_threshold = round(.threshold,3)) %>%
  mutate(diff_tprfpr = tpr - fpr) %>%
  slice_max(diff_tprfpr,n=1, with_ties = FALSE) %>%
  dplyr::select(fpr,tpr,score_threshold,ks = diff_tprfpr)

scored_test %>%
 roc_curve(loan_status, .pred_1) %>%
  autoplot() +
  geom_vline(aes(xintercept=0.03, color="red")) +
  labs(title="ROC operating at 3% FPR")

scored_test %>%
  pr_curve(loan_status, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3),
    f1 = round(2*precision*recall/(precision+recall),3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold),
            f1 = max(f1))

scored_test %>%
  ggplot(aes(.pred_1, fill=loan_status)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.1)) +
  labs(title="xgb score distribution")
```



# Explaination for final model
## explainer
```{r}
library(DALEXtra)

xgb_explainer <- explain_tidymodels(
  xgb_model_fit,
  data = dplyr::select(train, -loan_status),
  y = train$loan_status ,
  verbose = FALSE
)
```
## numeric Variable Funciton
```{r}
pdp_plotter <- function(variable){
  pdp_age <- model_profile(
  xgb_explainer,
  variables = variable
)
  
pdp_plot <- as_tibble(pdp_age$agr_profiles) %>%
  mutate(`_label_` = str_remove(`_label_`, "workflow_")) %>%
  ggplot(aes(`_x_`, `_yhat_`, color = `_label_`)) +
  geom_line(size = 1.2, alpha = 0.8) +
  labs(
    x = variable,
     y = " Average prediction Impact ",
    color = NULL,
    title = "Partial Dependence Profile Plot:",
    subtitle = variable
  )
print(pdp_plot)
}

numeric_vars <- c("last_pymnt_d_yr","last_pymnt_amnt","last_credit_pull_d_yr","member_id","last_credit_pull_d_mth","term","last_pymnt_d_mth","total_rec_late_fee","installment","issue_d_mth")

for (var in numeric_vars){
  pdp_plotter(var)
}


```
## Local Explainations
```{r}
model_features<-c(
  "last_pymnt_d_yr","last_pymnt_amnt","last_credit_pull_d_yr","member_id","last_credit_pull_d_mth","term","last_pymnt_d_mth","total_rec_late_fee","installment","issue_d_mth",'id', 'loan_amnt', 'funded_amnt', 'funded_amnt_inv', 'int_rate', 'grade', 'sub_grade', 'emp_length', 'home_ownership', 'annual_inc', 'verification_status', 'pymnt_plan', 'desc', 'purpose', 'zip_code', 'dti', 'delinq_2yrs', 'fico_range_low', 'fico_range_high', 'inq_last_6mths', 'open_acc', 'pub_rec', 'revol_bal', 'revol_util', 'total_acc', 'out_prncp', 'out_prncp_inv', 'application_type', 'acc_now_delinq', 'delinq_amnt', 'pub_rec_bankruptcies', 'tax_liens', 'earliest_cr_line_mth', 'earliest_cr_line_yr', 'issue_d_yr', 'addr_state_freq','sentiment_desc','title_freq', 'issue_d', 'earliest_cr_line', 'last_pymnt_d', 'last_credit_pull_d')


xgb_explainer <- explain_tidymodels(
    xgb_final_fit,   # fitted workflow object 
    data = train,    # original training data
    y = train$loan_status, # predicted outcome 
    label = "xgboost",
    verbose = FALSE)



explain_prediction <- function(record){
  record = record %>% dplyr::select(all_of(model_features))
  # run the explainer 
  xgb_breakdown <- predict_parts(explainer = xgb_explainer, 
                                 new_observation = record) %>% as_tibble()
  prediction_prob <- predict(xgb_model_fit,
        record,
        type="prob") %>% pull()
  # plot the explainer 
  p1 <- xgb_breakdown %>%
      filter(variable != "prediction")%>%
    dplyr::select(-desc)%>%
      mutate(contribution = round(contribution,3)) %>%
    filter(contribution!=0)%>%
      ggplot(aes(y=reorder(variable,position),x=contribution, fill=sign)) +
      geom_col() + 
      geom_text(aes(label=contribution), 
                position=position_dodge(width=0.7),
                vjust=0.5,)+
        labs(title = "DALEX explainations",
          subtitle = paste("predicted:",as.character(round(prediction_prob,3))),
                          x="contribution",
                          y="features")
  print(p1)
}

top_10_tp <- scored_test %>%
  filter(.pred_class == loan_status) %>%
  slice_max(.pred_1,n=10)

top_10_fp <- scored_test %>%
  filter(.pred_class != loan_status) %>%
   filter(loan_status == 0 ) %>%
  slice_max(.pred_1,n=10)

top_10_fn <- scored_test %>%
  filter(.pred_class != loan_status ) %>%
  filter(loan_status == 1 ) %>%
  slice_max(.pred_1,n=10)

# repeat for FP and FN 
for (row in 1:nrow(top_10_tp)) {
    s_record <- top_10_tp[row,]
    explain_prediction(s_record)                    
} 

for (row in 1:nrow(top_10_fp)) {
    s_record <- top_10_fp[row,]
    explain_prediction(s_record)
} 

for (row in 1:nrow(top_10_fn)) {
    s_record <- top_10_fn[row,]
    explain_prediction(s_record)
} 

```


# Anomaly Detection
## Train IsolationForest
```{r}
bake_loan<-bake(loan_recipe%>%prep(),loan)
iso_forest <- isolationForest$new(
  sample_size = 256,
  num_trees = 500,
  max_depth = ceiling(log2(256)))


iso_forest$fit(bake_loan)
```
## Pick average_depth
```{r}
pred_train <- iso_forest$predict(bake_loan)

pred_train %>%
  ggplot(aes(average_depth)) +
  geom_histogram(bins=20) + 
  geom_vline(xintercept = 7.2, linetype="dotted", 
                color = "blue", size=1.5) + 
  labs(title="Isolation Forest Average Tree Depth")

pred_train %>%
  ggplot(aes(anomaly_score)) +
  geom_histogram(bins=20) + 
  geom_vline(xintercept = 0.615, linetype="dotted", 
                color = "blue", size=1.5) + 
  labs(title="Isolation Forest Anomaly Score Above 0.615")


```
## Global Level Interpretation 
```{r}
train_pred <- bind_cols(iso_forest$predict(bake_loan),bake_loan) %>%
  mutate(anomaly = as.factor(if_else(average_depth <= 7.2, "Anomaly","Normal")))

train_pred %>%
  arrange(average_depth) %>%
  count(anomaly)

```
## Fit a Tree 
```{r}
fmla <- as.formula(paste("anomaly ~ ", paste(bake_loan %>% colnames(), collapse= "+")))

outlier_tree <- decision_tree(min_n=2, tree_depth=3, cost_complexity = .01) %>%
  set_mode("classification") %>%
  set_engine("rpart") %>%
  fit(fmla, data=train_pred)

outlier_tree$fit
```
## Plot the Tree
```{r}
library(rpart.plot) # -- plotting decision trees 

rpart.plot(outlier_tree$fit,clip.right.labs = FALSE, branch = .3, under = TRUE, roundint=FALSE, extra=3)

```
## Global Anomaly Rules 
```{r}
anomaly_rules <- rpart.rules(outlier_tree$fit,roundint=FALSE, extra = 4, cover = TRUE, clip.facs = TRUE) %>% clean_names() %>%
  #filter(anomaly=="Anomaly") %>%
  mutate(rule = "IF") 


rule_cols <- anomaly_rules %>% dplyr::select(starts_with("x_")) %>% colnames()

for (col in rule_cols){
anomaly_rules <- anomaly_rules %>%
    mutate(rule = paste(rule, !!as.name(col)))
}

anomaly_rules %>%
  as.data.frame() %>%
  filter(anomaly == "Anomaly") %>%
  mutate(rule = paste(rule, " THEN ", anomaly )) %>%
  mutate(rule = paste(rule," coverage ", cover)) %>%
  dplyr::select( rule)

anomaly_rules %>%
  as.data.frame() %>%
  filter(anomaly == "Normal") %>%
  mutate(rule = paste(rule, " THEN ", anomaly )) %>%
  mutate(rule = paste(rule," coverage ", cover)) %>%
  dplyr::select( rule)
```
## Show anomaly
```{r}

pred_train <- bind_cols(iso_forest$predict(bake_loan),
                        bake_loan)


pred_train %>%
  arrange(desc(anomaly_score) ) %>%
  filter(average_depth <= 7.2)%>%
  dplyr::select(id,average_depth,anomaly_score,inq_last_6mths,sentiment_desc,total_acc)
```








# kaggle
```{r}
kaggle <- read_csv("loan_holdout.csv") %>%
  clean_names() 
head(kaggle)
```

## Data Munging
```{r}
kaggle %>%
  dplyr::select(-mths_since_last_record,-mths_since_last_delinq,-next_pymnt_d,-url,-emp_title)%>%
  mutate(term = as.numeric(chartr('months','      ',term))) %>%
  mutate(int_rate = as.numeric(chartr('%',' ',int_rate))) %>%
  mutate(sub_grade = as.numeric(substr(sub_grade,2,3))) %>%
  mutate(zip_code = as.numeric(chartr('xx','  ',zip_code))) %>%
  mutate(revol_util = as.numeric(chartr('%',' ',revol_util))) %>%
  mutate(earliest_cr_line_mth = match(tolower(substr(earliest_cr_line,1,3)), tolower(month.abb))) %>%
  mutate(earliest_cr_line_yr = as.numeric(substr(earliest_cr_line,5,9))) %>%
  mutate(issue_d_mth = match(tolower(substr(issue_d,1,3)), tolower(month.abb))) %>%
  mutate(issue_d_yr = as.numeric(substr(issue_d,5,9))) %>%
  mutate(last_pymnt_d_mth = match(tolower(substr(last_pymnt_d,1,3)), tolower(month.abb))) %>%
  mutate(last_pymnt_d_yr = as.numeric(substr(last_pymnt_d,5,9))) %>%
  mutate(last_credit_pull_d_mth = match(tolower(substr(last_credit_pull_d,1,3)), tolower(month.abb))) %>%
  mutate(last_credit_pull_d_yr = as.numeric(substr(last_credit_pull_d,5,9))) %>%
  dplyr::select(-issue_d,-earliest_cr_line,-last_pymnt_d,-last_credit_pull_d,
                -collections_12_mths_ex_med,-policy_code,-chargeoff_within_12_mths)->kaggle

kaggle <- kaggle %>%
  left_join(addr_state_freq)%>%
  dplyr::select(-addr_state) ->kaggle

kaggle %>%
  mutate_if(is.character,as_factor) %>%
  mutate(desc=as.character(desc))->kaggle

kaggle <- kaggle %>% 
  left_join(sent_desc)

kaggle <- kaggle %>%
  left_join(title_freq)
kaggle %>%
  dplyr::select(-title) ->kaggle

kaggle%>%
  mutate(issue_d=issue_d_yr+issue_d_mth/12,
         earliest_cr_line=earliest_cr_line_yr+earliest_cr_line_mth/12,
         last_pymnt_d=last_pymnt_d_yr+last_pymnt_d_mth/12,
         last_credit_pull_d=last_credit_pull_d_yr+last_credit_pull_d_mth/12)->kaggle

```


```{r}
predict(xgb_final_fit, kaggle, type="prob") %>%
  bind_cols(kaggle) %>%dplyr::select(id,.pred_1)%>%rename(loan_status=.pred_1)->kaggle_perdict

write.csv(kaggle_perdict,'kaggle15.csv')
```

