---
title: "Churn prediction by Yinan Guo"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(vip) 
library(fastshap) 
library(rpart.plot) 
library(MASS)
library(caret)
library(rpart)
library(glmnet)
library(reshape2)
```

# Import training data set
```{r}
churn <- read_csv("Churn_training.csv") %>% clean_names()
head(churn)
```

# Explore data
## skim
```{r}
churn%>%skim_without_charts()
```

## Make factors
```{r}
churn <- churn %>%
  mutate(churn =as_factor(churn))%>%
  mutate_if(is.character,as_factor)
```

## Explore target
```{r}
churn %>%
  count(churn) %>%
  mutate(pct = n/sum(n)) -> churn_rate

churn_rate %>%
  ggplot(aes(x=churn, y=pct)) +
  geom_col() +
  geom_text(aes(label=pct) ,color="red") + 
  labs(title="Churn Rate")
```

## Explore numerics
```{r}
# Explore numeric
ggplot(churn, aes(x=monthly_minutes,y=churn,group=churn)) + geom_boxplot() +labs(title = "Monthly minutes ")
ggplot(churn, aes(x=customer_service_calls,y=churn,group=churn)) + geom_boxplot() +labs(title = "Customer Service Calls")
ggplot(churn, aes(x=streaming_minutes,y=churn,group=churn)) + geom_boxplot() +labs(title = "Streaming Munites")
ggplot(churn, aes(x=total_billed,y=churn,group=churn)) + geom_boxplot() +labs(title = "Total billed")
ggplot(churn, aes(x=prev_balance,y=churn,group=churn)) + geom_boxplot() +labs(title = "Prev Balance")
ggplot(churn, aes(x=late_payments,y=churn,group=churn)) + geom_boxplot() +labs(title = "Late Payments")
ggplot(churn, aes(x=phone_area_code,y=churn,group=churn)) + geom_boxplot() +labs(title = "Phone Area Code")
ggplot(churn, aes(x=number_phones,y=churn,group=churn)) + geom_boxplot() +labs(title = "Number Phones")

```

## Explore characters
### barplot
```{r}
bar_plot <- function(a,colname){
    churn %>%
    ggplot(., aes(a)) + 
    geom_bar(aes(fill = churn), position = "fill") +
    labs(title = paste("Barplot for",colname), x = colname, y="percent")+
    geom_hline(yintercept = 0.05)+
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5))
}
bar_plot(churn$email_domain,"Email Domain")
bar_plot(churn$phone_model,"Phone Model")
bar_plot(churn$partner,"Partner")
bar_plot(churn$phone_service,"Phone Service")
bar_plot(churn$streaming_plan,"Streaming Plan")
bar_plot(churn$multiple_lines,"Multiple Lines")
bar_plot(churn$mobile_hotspot,"Mobile Hotspot")
bar_plot(churn$wifi_calling_text,"Wifi Calling Text")
bar_plot(churn$online_backup,"Online Backup")
bar_plot(churn$device_protection,"device_protection")
bar_plot(churn$contract_code,"Contract Code")
bar_plot(churn$currency_code,"Currency Code")
bar_plot(churn$maling_code,"Maling Code")
bar_plot(churn$paperless_billing,"Paperless Billing")
bar_plot(churn$payment_method,"Payment Method")
bar_plot(churn$gender,"Gender")
bar_plot(churn$network_speed,"Network Speed")
```

### chart
```{r}
category_eval <- function(col){
  
churn%>% 
  group_by(churn) %>%
  count(!!as.name(col)) %>%
  pivot_wider(values_from=c(n), names_from = churn) %>%
      mutate(pct_1 = `1`/ (`0` + `1`),
             pct_0 = 1 - pct_1) %>%
    arrange(desc(pct_1)) %>%
    print()
}

for (col in colnames(churn %>% select_if(is.factor))){
  print(col)
  category_eval(col)
}
```

## Explore correlation
```{r}
library(corrplot)
cormat <- cor(select_if(churn%>%na.omit(),is.numeric))
round(cormat, 2)
corrplot(cormat)
```

## mutate date into days
```{r}
churn %>%
  mutate(regday=as.numeric(Sys.Date()-customer_reg_date))->churn
```

# Tain Test Split 
```{r}
set.seed(123)

train_test_spit<- initial_split(churn, prop = 0.7, strata = churn)

train <- training(train_test_spit)
test  <- testing(train_test_spit)

sprintf("Train PCT : %1.2f%%", nrow(train)/ nrow(churn) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(test)/ nrow(churn) * 100)
```

# Logistic Models
## Standard Logistic Model 
```{r}
churn_recipe <- recipe(churn ~ ., 
                      data = train) %>%
  step_rm(customer_id,billing_address,billing_city,billing_postal,customer_reg_date,gender) %>%
  step_novel(all_nominal_predictors()) %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())

logistic_spec <- logistic_reg() %>%
  set_mode("classification") %>%
  set_engine("glm")

logistic_full_wf <- workflow() %>%
  add_recipe(churn_recipe) %>%
  add_model(logistic_spec) %>%
  fit(train)

logistic_full_wf %>%
  pull_workflow_fit() %>%
  tidy() %>%
  mutate(across(is.numeric,round,3))

logistic_full_wf %>%
  pull_workflow_fit() %>%
  vip()

options(yardstick.event_first = FALSE)
predict(logistic_full_wf, train, type="prob") %>%
  bind_cols(predict(logistic_full_wf, train, type="class")) %>%
  bind_cols(train) ->logistic_train
logistic_train %>%
  metrics(churn, estimate = .pred_class, .pred_1)

predict(logistic_full_wf, test, type="prob") %>%
  bind_cols(predict(logistic_full_wf, test, type="class")) %>%
  bind_cols(test) -> logistic_test 

logistic_test %>%
  metrics(churn, estimate = .pred_class, .pred_1)
```

## Recuded Logistic
```{r}
churn_recipe1 <- recipe(churn ~ ., 
                      data = train) %>%
  step_rm(customer_id,billing_address,billing_city,billing_postal,customer_reg_date,
          senior_citizen,customer_service_calls,gender,
          partner,multiple_lines,wifi_calling_text) %>%
  step_novel(all_nominal_predictors()) %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())

logistic_r_wf <- workflow() %>%
  add_recipe(churn_recipe1) %>%
  add_model(logistic_spec) %>%
  fit(train)

logistic_r_wf %>%
  pull_workflow_fit() %>%
  tidy() %>%
  mutate(across(is.numeric,round,3))

logistic_r_wf %>%
  pull_workflow_fit() %>%
  vip()

options(yardstick.event_first = FALSE)
predict(logistic_r_wf, train, type="prob") %>%
  bind_cols(predict(logistic_r_wf, train, type="class")) %>%
  bind_cols(train)  ->logistic_r_train
logistic_r_train %>%
  metrics(churn, estimate = .pred_class, .pred_1)

predict(logistic_r_wf, test, type="prob") %>%
  bind_cols(predict(logistic_r_wf, test, type="class")) %>%
  bind_cols(test) -> logistic_r_test 
logistic_r_test %>%
  metrics(churn, estimate = .pred_class, .pred_1)
```

## Compare models 
```{r}
logistic_test %>%
  mutate(name="full logistic") %>%
  bind_rows(logistic_r_test %>% mutate(name="reduced logistic")) %>%
  group_by(name) %>%
  metrics(churn, estimate = .pred_class, .pred_1) %>%
  pivot_wider(id_cols=name, names_from=.metric,values_from=.estimate)%>%
  arrange(desc(roc_auc))
```

## Confusion Matrix
```{r}
logistic_test %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix full logistic test")

logistic_r_test %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix reduced logistic test")

logistic_train %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix full logistic train")

logistic_r_train %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="confusion matrix reduced logistic train")
```

## recall, precision, threshold
```{r}
logistic_test %>%
  pr_curve(churn, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))

logistic_r_test %>%
  pr_curve(churn, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))
```

## Score distribution
```{r}
logistic_test %>%
  #filter(churn==1) %>%
  ggplot(aes(.pred_1, fill=churn)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.282)) +
  labs(title="full logistic score distribution")

logistic_r_test %>%
  #filter(churn==1) %>%
  ggplot(aes(.pred_1, fill=churn)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.269)) +
  labs(title="reduced logistic score distribution")


```

## change threshold
```{r}
logistic_r_test %>%
  mutate(predict_class = as.factor(if_else(.pred_1 >=0.269,1,0))) %>%
  conf_mat(churn, estimate = predict_class) %>%
  autoplot(type = "heatmap") +
  labs(title="confusion matrix threshold >= 0.269")
```

## ROC Curve
```{r}
bind_rows(logistic_test %>%
  mutate(model = "full logistic test"), 
logistic_train %>%
  mutate(model = "full logistic train"),
logistic_r_test %>%mutate(model="reduced logistic test"),
logistic_r_train %>% mutate(model="reduced logistic train")) %>%
  group_by(model) %>%
  roc_curve(churn, .pred_1) %>%
  autoplot() +
  labs(title = "ROC chart")
```

# Decision trees
# Default tree
```{r}
tree_spec <- decision_tree(c=0) %>%
  set_mode("classification") %>%
  set_engine("rpart")

tree_wf1 <- workflow() %>%
  add_recipe(churn_recipe) %>%
  add_model(tree_spec) %>%
  fit(train)

tree_wf1 %>%
  pull_workflow_fit() %>%
  vip()

# -- deal w. the first event issue -- # 
options(yardstick.event_first = FALSE)
predict(tree_wf1, train, type="prob") %>%
  bind_cols(predict(tree_wf1, train, type="class")) %>%
  bind_cols(train) ->tree_train1
tree_train1%>%
  metrics(churn, estimate = .pred_class, .pred_1)

predict(tree_wf1, test, type="prob") %>%
  bind_cols(predict(tree_wf1, test, type="class")) %>%
  bind_cols(test) -> tree_test1

tree_test1 %>%
  metrics(churn, estimate = .pred_class, .pred_1)
```

## calculate cp
```{r}
bake_train <- bake(churn_recipe%>%prep(), new_data = train)
set.seed(123)
churn_tree1 <- train(churn ~ .,
                   data = bake_train, 
                  method="rpart",
                  trControl = trainControl("cv", number = 40),
                  tuneLength = 20)
plot(churn_tree1)
churn_tree1$bestTune
```

## Best cp tree
```{r}
tree_spec2 <- decision_tree(c=0.001609,tree_depth=10,min_n=2) %>%
  set_mode("classification") %>%
  set_engine("rpart")

tree_wf2 <- workflow() %>%
  add_recipe(churn_recipe) %>%
  add_model(tree_spec2) %>%
  fit(train)

tree_wf2 %>%
  pull_workflow_fit() %>%
  vip()

# -- deal w. the first event issue -- # 
options(yardstick.event_first = FALSE)
predict(tree_wf2, train, type="prob") %>%
  bind_cols(predict(tree_wf2, train, type="class")) %>%
  bind_cols(train) ->tree_train2
tree_train2%>%
  metrics(churn, estimate = .pred_class, .pred_1)

predict(tree_wf2, test, type="prob") %>%
  bind_cols(predict(tree_wf2, test, type="class")) %>%
  bind_cols(test) -> tree_test2
tree_test2 %>%
  metrics(churn, estimate = .pred_class, .pred_1)
```

## Compare models
```{r}
tree_test1 %>%
  mutate(name="Default tree") %>%
  bind_rows(tree_test2 %>% mutate(name="Best complexity tree")) %>%
  group_by(name) %>%
  metrics(churn, estimate = .pred_class, .pred_1) %>%
  pivot_wider(id_cols=name, names_from=.metric,values_from=.estimate) %>%
  arrange(desc(roc_auc))
```

## Confusion Matrix
```{r}
tree_test1 %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="default tree test")

tree_test2 %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="best complexity tree test")

tree_train1 %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="default tree train")

tree_train2 %>%
  conf_mat(churn, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + labs(title="best complexity tree train")
```

## ROC Curve
```{r}
bind_rows(tree_test1 %>%
  mutate(model = "default tree test"), 
tree_train1 %>%
  mutate(model = "default tree train"),
tree_test2 %>%mutate(model="reduced logistic test"),
tree_train2 %>% mutate(model="reduced logistic train")) %>%
  group_by(model) %>%
  roc_curve(churn, .pred_1) %>%
  autoplot() +
  labs(title = "ROC chart")
```

## recall, precision, threshold
```{r}
tree_test1 %>%
  pr_curve(churn, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))
tree_test2 %>%
  pr_curve(churn, .pred_1) %>%
  mutate(
    recall = round(recall, 2),
    .threshold = round(.threshold, 3),
    precision = round(precision, 3)
  ) %>%
  group_by(recall) %>%
  summarise(precision = max(precision),
            .threshold = min(.threshold))
```

## Score distribution
```{r}
tree_test1 %>%
  #filter(churn==1) %>%
  ggplot(aes(.pred_1, fill=churn)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.058)) +
  labs(title="default tree score distribution")

tree_test2 %>%
  #filter(churn==1) %>%
  ggplot(aes(.pred_1, fill=churn)) +
  geom_histogram(bins=100) +
  xlim(0, 1) +
  geom_vline(aes(xintercept=0.119)) +
  labs(title="best complexity tree score distribution")
```


